Title Goes Here

### 2021-04-08 analysis/paper/shceof.Rmd
Pone ecuación en el plot y agrega medidas del fit en el txto
Closes #16

```{r enso-phase, fig.cap = "Relationship between ENSO and phase of cEOF2 for the period 1979 -- 2019. Colours denote years with magnitude of cEOF2 greater or smaller than the 50th percentile. Black line is the fit ONI ~ cos(phase) + sin(phase) computed by OLS weighted by the magnitude of cEOF2.", fig.width = width_column, fig.height=3}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit()

@@

@@

  scale_color_brewer(NULL, palette = "Dark2",
                     labels = c("TRUE"  = "|EOF2| > 50%",
                                "FALSE" = "|EOF2| < 50%")) +
  scale_shape_manual(NULL, values = c("TRUE" = 20,
                                      "FALSE" = 18),
                     labels = c("TRUE"  = "|EOF2| > 50%",
                                "FALSE" = "|EOF2| < 50%")) +
  scale_y_continuous("ONI") +
  scale_x_continuous("Phase",
                     breaks = seq(-pi, pi, by = 45*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  coord_cartesian(xlim = c(-pi, pi)) +
  panel_background
```

To better explore the relationship between tropical forcing and phase of the cEOF2, Figure\ \@ref(fig:enso-phase) plots the ONI index and the phase of the cEOF2 for each SON trimester between 1979 and 2019, highlighting years in which the magnitude of cEOF2 is above the median. In years with positive ONI, the phase of the cEOF2 is mostly around +90º (corresponding with positive imaginary part) and vice versa. In years with near neutral ENSO, the phase of the cEOF2 is much more variable. The black line in Figure\ \@ref(fig:enso-phase) is an harmonical fit of the relationship between ONI and cEOF2 phase. The $r^2$ corresponding to the fit is `r signif(summary(cos_model)$r.squared, 2)`, with p-value `r report_p(coef(summary(cos_model))[4])`, indicating a quasi-sinusoidal relation between these two variables.

@@

(ref:australia-pp-phase-cap) Same as Figure\ \@ref(fig:enso-phase) but for precipitacion anomalies averaged between  `r knitr::combine_words(LonLabel(box_australia$lon))` and `r knitr::combine_words(LatLabel(box_australia$lat))` (box shown in Figure\ \@ref(fig:pp-oceania) column b).

@@

@@

  scale_color_brewer(NULL, palette = "Dark2",
                     labels = c("TRUE"  = "|EOF2| > 50%",
                                "FALSE" = "|EOF2| < 50%")) +
  scale_shape_manual(NULL, values = c("TRUE" = 20,
                                      "FALSE" = 18),
                     labels = c("TRUE"  = "|EOF2| > 50%",
                                "FALSE" = "|EOF2| < 50%")) +
  scale_y_continuous("Precipitation anomaly") +
  scale_x_continuous("Phase",
                     breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  coord_cartesian(xlim = c(-pi, pi)) +
  panel_background
```

To better understand the relationship between precipitation anomalies over Western Australia and the cEOF2, Figure\ \@ref(fig:australia-pp-phase) plots precipitation anomalies averaged over Eastern Australia (box in Figure\ \@ref(fig:pp-oceania)) as a function of cEOF2 phase. The black line is the equation $\mathrm{Precipitation} = `r signif(coef(model)[1], 2)`\sin{(\mathrm{phase})} + `r signif(coef(model)[2], 2)`\cos{(\mathrm{phase})}$, whose coefficient we fitted by weighted least squares using the magnitude of the EFO2 as weight. 
\textcolor{red}{(Como te puse antes la ecuación como mucho la pondría sobre la figura)}
The $r^2$ corresponding to the fit is `r signif(summary(model)$r.squared, 2)`, with p-value `r report_p(p_value)`. There is a strong relationship between the phase of the cEOF2 and precipitation anomalies in Eastern Australia, but Real and Imaginary phases we chose are not exactly aligned with the direction that maximises this relationship. If one were to perform a more in-depth analysis of the relationship between this pattern and precipitation in this region, one would conceivably align one of the axis (either the Real or the Imaginary) to the one that maximises it.

>>>>

```{r enso-phase, fig.cap = "Relationship between ENSO and phase of cEOF2 for the period 1979 -- 2019. Colours denote years with magnitude of cEOF2 greater or smaller than the 50th percentile. Black line is the fit ONI ~ sin(phase) computed by OLS weighted by the magnitude of cEOF2.", fig.width = width_column, fig.height=3}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit()

@@

equation <- latex2exp::TeX(paste0("$\\mathrm{ONI} = ", 
        signif(coef(cos_model)[1], 2), "\\sin{(\\mathrm{phase})}$"),
        output = "text")

@@

  scale_color_brewer(NULL, palette = "Dark2",
                     labels = c("TRUE"  = "|EOF2| > 50%",
                                "FALSE" = "|EOF2| < 50%")) +
  scale_shape_manual(NULL, values = c("TRUE" = 20,
                                      "FALSE" = 18),
                     labels = c("TRUE"  = "|EOF2| > 50%",
                                "FALSE" = "|EOF2| < 50%")) +
  scale_y_continuous("ONI") +
  scale_x_continuous("Phase",
                     breaks = seq(-pi, pi, by = 45*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  coord_cartesian(xlim = c(-pi, pi)) +
  labs(subtitle = equation) +
  panel_background
```

To better explore the relationship between tropical forcing and phase of the cEOF2, Figure\ \@ref(fig:enso-phase) plots the ONI index and the phase of the cEOF2 for each SON trimester between 1979 and 2019, highlighting years in which the magnitude of cEOF2 is above the median. In years with positive ONI, the phase of the cEOF2 is mostly around +90º (corresponding with positive imaginary part) and vice versa. In years with near neutral ENSO, the phase of the cEOF2 is much more variable. The black line in Figure\ \@ref(fig:enso-phase) is a sinusoidal fit of the relationship between ONI and cEOF2 phase. The $r^2$ corresponding to the fit is `r signif(summary(cos_model)$r.squared, 2)`, with p-value `r report_p(coef(summary(cos_model))[4])`, indicating a quasi-sinusoidal relation between these two variables.

@@

(ref:australia-pp-phase-cap) Same as Figure\ \@ref(fig:enso-phase) but for precipitation anomalies averaged between  `r knitr::combine_words(LonLabel(box_australia$lon))` and `r knitr::combine_words(LatLabel(box_australia$lat))` (box shown in Figure\ \@ref(fig:pp-oceania) column b).

@@

equation <- latex2exp::TeX(paste0("$\\mathrm{Precipitation} = ", 
        signif(coef(model)[1], 2), "\\sin{(\\mathrm{phase})} + ", 
        signif(coef(model)[2], 2), "\\cos{(\\mathrm{phase})}$"), output = "text")

@@

  # annotate("text", x = 90*pi/180, y = 1, label = equation, parse = TRUE) +

  scale_color_brewer(NULL, palette = "Dark2",
                     labels = c("TRUE"  = "|EOF2| > 50%",
                                "FALSE" = "|EOF2| < 50%")) +
  scale_shape_manual(NULL, values = c("TRUE" = 20,
                                      "FALSE" = 18),
                     labels = c("TRUE"  = "|EOF2| > 50%",
                                "FALSE" = "|EOF2| < 50%")) +
  scale_y_continuous("Precipitation anomaly") +
  scale_x_continuous("Phase",
                     breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  coord_cartesian(xlim = c(-pi, pi)) +
  labs(subtitle = equation) +
  panel_background
```

To better understand the relationship between precipitation anomalies over Western Australia and the cEOF2, Figure\ \@ref(fig:australia-pp-phase) plots precipitation anomalies averaged over Eastern Australia (box in Figure\ \@ref(fig:pp-oceania)) as a function of cEOF2 phase. The black line shows a sinusoidal fit, whose coefficients we fitted by weighted least squares using the magnitude of the EFO2 as weight. The $r^2$ corresponding to the fit is `r signif(summary(model)$r.squared, 2)`, with p-value `r report_p(p_value)`. There is a strong relationship between the phase of the cEOF2 and precipitation anomalies in Eastern Australia, but Real and Imaginary phases we chose are not exactly aligned with the direction that maximises this relationship. If one were to perform a more in-depth analysis of the relationship between this pattern and precipitation in this region, one would conceivably align one of the axis (either the Real or the Imaginary) to the one that maximises it.
